---
title: "Preliminary Research Results"
author: "2024 FIRE198 Sustainability Analytics"
format: gfm
editor: visual
---

## Team Members

Lasya Muthyam, Sriya Peddinti

## Research Question

Effects of Plastic Bag Regulations on the Chesapeake Bay

## Background

Plastic bag bans have been implemented all around the Chesapeake Bay area in the past decade. Harford County in Maryland implemented a plastic bag ban in September 2017. Arlington and Fairfax counties in Virginia implemented a plastic bag tax in January 2022.

Seeley et al. (2020) and Shen et al. (2020) discussed how microplastics can impact ocean carbon sequestration and dissolved organic carbon levels.

Our research attempts to see what effects, if any, plastic bag bans and taxes have on dissolved carbon levels in the Chesapeake Bay.

## Data

The list of counties with plastic bag bans and taxes came from banthebag.com.

The list of counties and states came from the Chesapeake Bay Foundation website, which listed the counties that are considered to be a part of the Chesapeake Bay watershed.

Water quality data came from the EPA US Water Quality Portal. We looked at the dissolved carbon levels in a date range from 1-1-2010 to 3-1-2024 in the states/regions of New York, Pennsylvania, Maryland, Delaware, DC, Virginia, and West Virginia.

Clean data here.

```{r}

```

## Preliminary Results

Use your final dataset to visualize how the treatment variable impacts the outcome variable here:

The effects of the plastic bag ban on microplastic pollution in the Chesapeake Bay are unclear from our current results. We cannot conclude that the treatment variable, ban/tax, makes a significant impact on the outcome variable, average dissolved carbon levels.

```{r}

```
